\documentclass[12pt, letterpaper]{article}
\usepackage[margin=.5in]{geometry}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{adjustbox}

\title{
	GV101 Intro to PolSci\\
	\large{Professor Simon Hix}\\
	\large{Short Answer Questions}\\ {Revision Document}
}
\author{Cedric Tan}
\date{April 2019}


\begin{document}
\maketitle
\abstract{This document covers the 8 topics that could be asked in the short answer section of the GV101 paper. There is a section dedicated to each topic and each section will begin with the definitions related to the subject then go into moderate depth on the relevant material before concluding with a half page summary on the topic. 

Note that this is a collation of readings, lecture notes and further supplementary material that can be found online. All work is not necessarily authentic and has been modified by me to fit my needs for this course}

\newpage
\tableofcontents
\newpage


\section{The Modernisation Theory of Democratisation}

\textbf{Definition:} Classic modernisation theory argues that countries are more likely to become democratic and stay democratic as they develop economically $\rightarrow$ this theory is more prevalent in high-income countries. (Clark, Golder and Golder 2017)

\subsection{Key Ideas}
Modernisation theory argues that all societies pass through the same historical stages of economic development. Those who are not democratic are simply \textbf{labelled as underdeveloped.} Rostow (1960) and Gerschenkron (1962) believed that African, Asian and Latin American countries were simply underdeveloped versions of countries in Europe.

These 'primitive' countries were characterised by large agricultural sectors and small industrial and service sectors. Eventually these countries are supposed to modernise by expanding their industrial and service sectors while reducing their emphasis on the level of agriculture.

\subsection{The Political Science Approach}
Helmed by Lipset (1959, 1960), modern societies, he says, need an appropriate type of government. Przeworski (1998, 2000) states that dictatorships and other types of government are replaced by democracies because:
\begin{itemize}
	\item Social structure becomes more complex
	\item New groups emerge and organise along various lines
	\item Labour requires active cooperation by employees hence the system can no longer be \textbf{command driven.}
	\item Dictatorships lose control and effectiveness as:
		\begin{itemize}
			\item Technological change endows private information and autonomy
			\item Civil democratic society tends to emerge as a result
		\end{itemize}
\end{itemize}
Hence for Przeworski, modernisation theory highlights the idea that as a country develops economically, they will also democratise due to the plurality of opinions that are formed along with the independence individuals develop with their new information.

Lipset also argues that higher income countries will also tend to maintain their democratic status as:
\begin{itemize}
	\item Sustaining the democracy is done through the people and their interests
	\item Interest in democracy is a main concern for these people and it persists as long as Przeworksi's reasons of diverse social structure remain
\end{itemize}

\subsection{Summary of Modernisation Theory}
Summary to be independently written for extended revision.

\newpage
\section{The Selectorate Theory for Non-Democratic Regimes}

\textbf{Definition:} Selectorate theory helps to explain why we observe tremendous variation in the economic performance of dictatorships. Rather than categorise governments as either democratic or dictatorial, selectorate theory characterises all governments by their location in a two-dimensional institutional space:
\begin{itemize}
	\item One dimension is the size of the selectorate: those with a say in selecting the leader
	\item The second is the size of the winning coalition: those in the selectorate whose support is essential for the leader to stay in office
\end{itemize}

\begin{center}
	$W = Winning\;Coalition$ and $S = Selectorate$
\end{center}

\subsection{Types of Selectorate and Winning Coalition}
Here are various forms of the relationship between Selectorate and Winning Coalition:
\begin{itemize}
	\item Large W and Large S
		\begin{itemize}
			\item Democracies usually
			\item Incentive to produce public goods
			\item Good government performance
			\item High levels of wealth, efficient governane and low rates of corruption and kleptocracy
		\end{itemize}
	\item Small W and Large S
		\begin{itemize}
			\item Personalist dictatorships and dominant party democracies
			\item Incentives to provide rewards to their relatively small winning coalition to stay in power
			\item Rigged voting hence the large selectorate but small winning coalition
			\item Poor government performance
			\item Low levels of wealth, inefficient governance with high levels of corruption and kleptocracy
		\end{itemize}
	\item Small W and Small S
		\begin{itemize}
			\item Monarchic and military dictatorships
			\item Produces middling government performance
			\item Moderate levels of corruption and kleptocracy
		\end{itemize}
\end{itemize}

The basic assumption of selectorate theory is that all political leaders are motivated by the desire to gain office. The competitive nature of politics forces all leaders to behave in this way. With that in mind, government performance is derivative of the ruling power making their selectorate happy. As shown above, personalistic or dominant party regimes will have lacking performance due to a small W that needs to be appeased. Dictatorships tend to focus on private goods to be given to their W as a result of this. For democracys, who have large Ws, public goods are the focus.

\subsection{Loyalty Norm}
The \textbf{Loyalty Norm} extends the idea of keeping the respective Ws happy. The Loyalty Norm is determined by $W/S$ which is effectively the probability that someone in the selectorate is in the winning coalition. \\

\textbf{Low Probability:}

There is less chance of a member of W defecting as the odds that they could form a new winning coalition is low, hence the loyalty here is high. Strong loyalty norm regimes tend to have a \textbf{greater chance in engaging in corruption and kleptocracy.} The amount to pay W is lower to keep their loyalty. \\

\textbf{High Probability:}

These is a higher chance of a member of W defecting as the odds that they could form a new winning coalition is high, hence the loyalty here is low. Low loyalty norm regimes tend to have a \textbf{lower chance in engaging in corruption and kleptocracy.} The amount required to pay W is higher to keep their loyalty. 

\begin{center}

We can recognise the probabilities by constructing a simple equation:

% insert equation and example from Clark, Golder and Golder
$R_L = Reward\;for\;Loyalty$ ;
$R_D = Reward\;for\;Defecting$ \\
$P_L = Probablity\;of\;Staying$ ;
$P_D = Probability\;of\;Defecting$ \\

\end{center}

\subsection{Size of the Winning Coalition}
The size of W can affect the eocnomic performance heavily through either investment into public or private goods. Leaders \textbf{always prefer to use private goods to satisfy the winning coalition rather than public goods} as it is easier and more effective. However, \textbf{increasing W should lead to more public good production.} This is simply because there is more people to please and public goods, since they are \textbf{non-rivalrous} \textit{(one person's consumption does not hinder another's ability to consume)} and \textbf{non-excludable} \textit{(one cannot prevent another from accessing this good entirely)}, appease everyone. Private goods, on the other hand, can only entertain a few people and so would not be able to maintain W.

\subsection{Summary of Selectorate Theory}
Summary to be independently written for extended revision.


\newpage
\section{Differences between Strategic and Expressive Voting}
\textbf{Definitions:}
\begin{itemize}
	\item \textbf{Strategic:} voting to produce an election outcome which is as close as possible to one's policy preferences (which may or may not mean voting for one's preferred party).
	\item \textbf{Expressive or Sincere:} voting on the basis of party attachment, political ideology or social group membership.
\end{itemize}

The two are often the same i.e. expressively voting is usually strategic voting as well. However, the interesting scenario is when one is \textbf{cross-pressured:} where voting one way means not voting the other way, \textbf{when the two styles of voting are mutually exclusive.}

Voters are assumed to be in a spatial model of politics where each voter has preferences about a range of policies. These preferences are \textbf{single peaked} which means that each voter has an \textbf{ideal point} in a single or multi-dimensional policy space.

\subsection{Expressive}
So as we noted before, expressive means:
\begin{itemize}
	\item Identification with a particular social cleavage, group or ideology
	\item People voting in line with this identity whether it be socio-economic class, ethnic or cultural associations
	\item Religion (Tilley 2015) may also have an effect on how one would vote
	\item Voting for whatever party feels \textbf{closest to me.}
\end{itemize}

Hence, voting expressively is pretty self-explanatory. It is because it is closest to your personal preference that you would vote in that way. If you vote purely in terms of expression, strategic voting does not matter whatsoever. When you are cross-pressured, you will choose the party closest to you.

\subsection{Strategic}
And for strategic we see that:
\begin{itemize}
	\item Citizens vote to try to influence the outcome of an election
	\item Overall policy outcomes that they attempt to influence would be closest to their ideal policy of all the likely outcomes
	\item Citizens attempt to influence outcomes whilst maintaining policy preferences close to their true preferences
\end{itemize}

\subsection{Why vote strategically?}
As opposed to expressive voting, the reasons for voting strategically can differ. Not that when we talked about cross-pressure, the expressive voter did not mind. However, for strategic voters, especially for those who support smaller parties, it is a serious problem as smaller party representatives are unlikely to be elected into office so strategic voters need to reconfigure their vote to influence a policy outcome.

There are two main reasons:
\begin{enumerate}
	\item \textbf{Local:} to influence the election outcome in a constituency. If the candidate a person most preferences has no chance of being elected, then vote for the closest candidate in preference from amongst the candidates who have a reasonable chance of winning. Example:
		\begin{itemize}
			\item British Elections
			\item First preference: Liberal Democrat
			\item \% Voting for Lib Dem as first preference: 56\%
			\item Likelihood is that the others split between Labour and Conservative
		\end{itemize}
	\item \textbf{National:} to influence the government formation and policy outcomes. If the party a person most prefers has no choice of influencing government formation or might form a coalition with a party which is further away from a person's preference, then vote for a party which is 'further away' but which will lead to a policy outcome closer to a person's ideal policy. Example:\\
	\begin{center}
		\begin{tikzpicture}
			\draw (-5,0) -- (5,0);
			\draw (-5, .2) -- (-5, -.2);
			\draw (5, .2) -- (5, -.2);
			\coordinate [label=below:$V_L$] (A) at (-3,-.2);
				\draw (-3, 0) -- (-3, -.2);
			\coordinate [label=below:$V_R$] (B) at (3,-.2);
				\draw (3, 0) -- (3, -.2);
			\coordinate [label=below:$E_C$] (C) at (0,-.2);
				\draw (0, 0) -- (0, -.2);
			\coordinate [label=below:$P_{CL}$] (D) at (-2,-.2);
				\draw (-2, 0) -- (-2, -.2);
			\coordinate [label=below:$P_{CR}$] (E) at (2,-.2);
				\draw (2, 0) -- (2, -.2);
			\coordinate [label=below:$P_{FL}$] (F) at (-4.5,-.2);
				\draw (-4.5, 0) -- (-4.5, -.2);
			\coordinate [label=below:$P_{FR}$] (G) at (4.5,-.2);
				\draw (4.5, 0) -- (4.5, -.2);
			\draw [->] (-3, .5) -- (-2, .5);
			\draw [->] (-3, -1) -- (-0.1, -1);
			\draw [->] (-3, -1.5) -- (-4.5, -1.5);
			\coordinate [label=above:${Original}$] (H) at (-2.5, .5);
			\coordinate [label=below:${New}$] (I) at (-1.5, -1);
			\coordinate [label=below:${Shift}$] (J) at (-3.75, -1.5);
			\draw [->] (3, .5) -- (2, .5);
			\draw [->] (3, -1) -- (0.1, -1);
			\draw [->] (3, -1.5) -- (4.5, -1.5);
			\coordinate [label=above:${Original}$] (K) at (2.5, .5);
			\coordinate [label=below:${New}$] (L) at (1.5, -1);
			\coordinate [label=below:${Shift}$] (M) at (3.75, -1.5);
		\end{tikzpicture}
	\end{center}
	
	Hence this result illustrates a shift in voting as a result of a delcared coalition that was further away from the voter's original preferences. Look at $V_L$ which is the left voter and $V_R$ which is the right voter, they would originally vote for the parties on the center-left and center-right respectively. However, after a coalition is announced between those two parties, the new distance from their true preferences is shifted all the way to the middle. They will subsequently shift their vote for the far left and far right parties, although further away, to see some compromise on policy preferences.
\end{enumerate}

\subsection{Summary of Expressive and Strategic Voting}


\newpage
\section{Down's Theory of Party Competition}
\textbf{Definition:} the simple Downsian model of two-party competition under plurality is generally characterised as predicting party convergence to the policy position espoused by the median voter.

\subsection{Assumption of the Downsian Model}
Below are a list of assumptions that Grofman (2004) thinks Downs has
\begin{enumerate}
	\item The system only has two political parties
	\item Single-round election for any office
	\item The election chooses a single candidate
	\item Elections take place within a single constituency
	\item The election is decided by a plurality vote
	\item Policies can be located a single left-right dimension
	\item Candidate policy positions are well defined
	\item Candidate policy positions are accurately estimated by each voter
	\item Voters look no further than the next decision
	\item Eligible voters go to the polls if the expected benefits of their vote's contribution to the election exceeds costs
	\item Voters care only about which candidate/party will enact policies closest to their preference location. If there are no policy differences, they are equally likely to support each and every candidate
	\item Parties/candidates care only about winning
	\item Parties/candidates look no further than the next election
	\item Candidates/parties accurately estimate the policy preferences of voters, or at minimum, they can identify the location of the median voter overall or in each party
	\item Candidates are part of a unified party team
\end{enumerate}

\subsection{The Downsian Model}
Assuming that there is a normal distribution of voters, we can model the vote distribution and subsequent convergence as shown below:
\begin{center}
	\begin{tikzpicture}
		\coordinate [label=left:$L$] (A) at (-5, 0);
		\coordinate (B) at (0, 6);
		\coordinate [label=below:$P_A$] (P_A) at (-3, 0);
		\coordinate [label=below:$P_B$] (P_B) at (0, 0);
		\coordinate [label=below:$Split$] (M_P) at (-1.5, 0);
		\coordinate (C) at (0, 0);
		\coordinate [label=right:$R$](D) at (5, 0);
		\draw [thick, dashed] (C) -- (B);
		\draw [thick](A) -- (D);
		\draw (M_P) -- (-1.5 , 5);
		\draw (A) .. controls (0, 7) .. (D);
		\coordinate [label=left:$Vote\;Share$] (Top) at (-5, 6);
		\draw [thick, ->](A) -- (Top);
	\end{tikzpicture}
\end{center}

From the picture above, we can see that the split vote share is given mostly to B as party B is on the median voter line. Hence, the voter base is split at the middle point between $P_A$ and $P_B$. For A to gain more voters, it needs to shift its policy position closer to B so that the split becomes more even. The Downsian model then predict that A will keep shifting until the party reaches the median voter like B. \textbf{It would be the same case if B started at a further right position and they both converged to the median voter. \textit{In game theory this would be called a Nash Equilibrium.}} This would end up looking as you would expect:
\begin{center}
	\begin{tikzpicture}
		\coordinate [label=left:$L$] (A) at (-5, 0);
		\coordinate (B) at (0, 6);
		\coordinate [label=below:$P_A$] (P_A) at (-3, 0);
		\draw [->] (-2.5, -.3) -- (-.9, -.3);
		\coordinate [label=below:$P_{AN}$] (P_AN) at (-0.3, 0);
		\coordinate [label=below:$P_B$] (P_B) at (0.3, 0);
		\coordinate (C) at (0, 0);
		\coordinate [label=right:$R$](D) at (5, 0);
		\draw [thick, dashed] (C) -- (B);
		\draw [thick](A) -- (D);
		\draw (A) .. controls (0, 7) .. (D);
		\coordinate [label=left:$Vote\;Share$] (Top) at (-5, 6);
		\draw [thick, ->] (A) -- (Top);
	\end{tikzpicture}
\end{center}

Thus, both parties have an equal share of the votes as shown by the diagram as $P_A$ shifts to $P_{AN}$.

\subsection{Reasons Why Parties Might Not Converge}
Below are some reasons that parties might not converge which you can be wary of:
\begin{itemize}
	\item Party positions are sticky due to the costs associated with moving
	\item There is threat of competition by flanks. Far right and Far left parties might arise as a result of Downsian convergence
	\item The reality of a multidimensional space, rather than the assumed single one, makes it more difficult to know where to move; it isn't always the median voter!
	\item Low turnout could happen since the fringes are not accounted for. It might be easier to mobilise your base than to capture new voters and persuade them to switch
	\item It is also dependent on how party leaders are chosen, whether it be the average voter, party officials or members which get to choose. This affects whether they are more extreme or moderate
 
\end{itemize}

\subsection{Summary of Downs Party Theory}

\newpage
\section{Olson's Theory of Collective Action}
\textbf{Definitions:} 
\begin{itemize}	
	\item \textbf{Collective Action:} refers to the pursuit of some objective by a group of individuals. Typically the objective is some form of public or private good.
	\item \textbf{Public Good:} a good which is non-excludable, whereby one cannot prevent other people from consuming the good once it is produced, and non-rivalrous, where one's consumption of the good does not affect another's ability to consume the same good.
	\item \textbf{Interest Group:} Any group which seeks to promote a particular policy or set of policies, and to organise to influence politics and policy-makers to achieve this policy.
	\item \textbf{Social Movement:} A large informal group of individuals and/or organisations which aim to promote a particular political or social issue, or promote/resist social change.

\end{itemize}

\subsection{The Logic of Collective Action}
Olson attempts to explain why some groups are more able to mobilise than others and, as a result, why public goods are likely to be undersupplied and private goods oversupplied. He presents us with a \textbf{collective action function} shown below:

\begin{center}
Each individual has a collective action function: \\
$R=(B \times P) - C$ \\
$R$ is the reward for participating in collective action \\
$B$ is the benefit of the good provided by a group \\
$P$ is the probability that the action of the individual makes a difference \\
$C$ is the cost of participation
\end{center}

\begin{center}
\textbf{Example 1: Student Demonstrations} \\
This example is to show the free rider problem we can face in collective action. \\
Taking $R=(B \times P) - C$ \\
$B$ would be cutting tuition fees \\
$P$ would be low as government probably won't change their stance \\
$C$ would be high as you're marching the whole day in the cold! \\
Hence we would have \\
$R=(B \times P) - C$ where $R<0$
\end{center}

Thus there's the free rider problem as individuals have little incentive to join a group which seeks a public good, as they all benefit if the protests are successful $\rightarrow$ there will be students protesting anyway. However, the costs for the individuals exceed the probable benefits.

\begin{center}
\textbf{Example 2: Farm Subsidies} \\
This example shows the power of organised interests \\
For this example, take 20 citizens in society with 4 out of the 20 being farmers. Each citizen \textbf{pays £1 in tax} and each farmer \textbf{receives a £5 subsidy.}

\begin{center}
\begin{tabular}{|c|c|c|c|}

	\hline
	& Tax & Subsidy & Outcome \\
	\hline
	\hline
	Farmers & £1 & £5 & +£4 \\
	\hline
	Individuals & £1 & £0 & -£1 \\
	\hline

\end{tabular}
\end{center}

Taking $R=(B \times P) - C$ \\
$B$ would be cutting tuition fees \\
$P$ would be low as government probably won't change their stance \\
$C$ would be high as you're marching the whole day in the cold! \\
Hence we would have \\
$R=(B \times P) - C$ where $R<0$
\end{center}

Each farmer has an incentive to organise to maintain subsidies while each taxpayer and consumer has no incentive to stop them. This is because farmers represent a concentrated interest seeking a \textbf{private good} while taxpayers represent a diffuse interest, \textbf{they are not aligned unless the tax is insanely high!} Hence farmers mobilise and more likely to get what they want.

\subsection{Summary of Olson's Theory of Collective Action}


\newpage
\section{Tsbelis' Theory of Veto Players and Agenda Setters}
\textbf{Definitions:}
\begin{itemize}
	\item \textbf{Veto Players:} actors whose agreement is required for a change of the status quo, they have the right to block a proposal e.g. the median member of parliament
	\item \textbf{Agenda Setters:} the right or ability to make a proposal or influence the importance placed on the topics of the public agenda e.g. the government
\end{itemize}

\subsection{Tsebelis Overview}
His view is based on the political system as the means for collective decisionmaking. Consequently, all actors in the system, whether they be voters, representatives or political parties, care about policy outcomes, either directly or indirectly $\rightarrow$ either because they have preferences over outcomes or because other things they like depend on policy outcomes.

Policy outcomes are the result of two factors:
\begin{itemize}
	\item The preferences of the actors involved
	\item Prevailing institutions in place
\end{itemize}
More importantly, understanding where the status quo is can affect the stability of policy.

Tsebelis believes that all political institutions are translated into a series of veto-players and agenda-setters. The number and location of veto players affects policy stability, or \textbf{how difficult it is to change the status quo.}

Further, the sequence in which veto players make their decisions affects the influence that these veto players have in the decisionmaking process. If you're an individual veto-player or in a collective group of veto-players, your decisionmaking process will also differ.

\subsection{Individual Veto Players}
We will focus on individual veto players and how they affect policy in simple Euclidean spatial models.

He has two main propositions:
\begin{itemize}
	\item \textbf{More Veto Players $\rightarrow$ Less Policy Change}
	
	Examples include: coalition governments, presidential systems, bicameralism, supreme court, central bank
	\item \textbf{Bigger Policy Distance between Veto Players $\rightarrow$ Less Policy Change}
	
	Examples include: Coalition between two ideologically similar vs between two ideologically different
\end{itemize}

It follows that a change in the status quo requires a unanimous decision of all veto players. 

If veto players are generated by the constitution, they are called \textbf{institutional veto players.} For example, the US Constitution specifies that legislation, to be enacted, requires the approval of the president, House of Representatives and the senate. All three parts of are seen to be veto players.

If veto players are generated by the political game, they are called \textbf{partisan veto players.} For example, it may be that inside the House of Representatives, different majorities exist, the majority party is the real (partisan) veto player even though the house itself is an institutional veto player.\\\\
Each individual veto player is represented by his/her ideal point in an \textit{n-dimensional} policy space. In addition, each veto player has \textit{circular indifference curves}, indifferent between alternatives that have the same distance from his ideal point.
Two more concepts are identified by Tsebelis:
\begin{itemize}
	\item \textbf{Winset of the Status Quo:} the set of policies that would beat the status quo in a pair-wise contest under whatever voting rules are being employed i.e. replace existing policy
	\item \textbf{Core:} the set of points with empty winset i.e. the points that cannot be defeated by any other point if we apply the decisionmaking rule. For example, the \textit{unanimity core} refers to the set of points that cannot be defeated if the decision is unanimous.
\end{itemize}
Thus, depending on these concepts and their relative significance or size, policy stability is affected. Policy stability is the difficult of effecting significant change in the status quo. If the winset is small, policy is more stable as there is less room to change. If the winset is large, policy tends to be more flexible as the area of indifference between the veto players are larger:
\begin{itemize}
	\item If the winset of the status quo exists, its size decreases or remains the same with the addition of new veto players.
	\item If there is a unanimity core, its size increases or remains the same with the addition of new veto players.

\end{itemize}

\textbf{Hence, the addition of a new veto player increases policy stability or leaves it the same.} Through Tsebelis' theory, we can see how much consesus and how many constraints we have and need on an agenda setter.

\subsection{Operationalising the Theory}
Below are examples of operationalising the theory.
\subsubsection{Majoritarian Goverment: Dictatorship of the Majority Party}
\textbf{Assumptions:}
\begin{itemize}
	\item $A$, $B$ and $C$ are in the Left Party and $D$ and $E$ are in the Right Party
	\item $B$ is the leader of the Left Party $\rightarrow$ \textbf{$B$ is the agenda setter}
	\item Following these, if there is party cohesion, B is the dictator
\end{itemize}

\begin{center}
\begin{tikzpicture}
\coordinate [label=left:$Left$] (Left) at (-5,0);
\coordinate [label=right:$Right$] (Right) at (5,0);
\coordinate [label=below:$A_L$] (A) at (-3.5,0);
\coordinate [label=below:$B_L$] (B) at (-2,0);
\coordinate [label=below:$C_L$] (C) at (0,0);
\coordinate [label=below:$D_R$] (D) at (1,0);
\coordinate [label=below:$E_R$] (E) at (3,0);
\coordinate [label=above:$SQ$] (SQ) at (2, 1);
\coordinate [label=above:$X$] (X) at (-2, 1);
\draw [->] (SQ) -- (2, 0);
\draw [->] (X) -- (B);
\draw (Left) -- (Right);

% Draw in status quo with chosen policy X

\end{tikzpicture}
\end{center}
This is because $A$, $B$ and $C$ control the majority in the government and any agenda put through by the agenda setter, $B$, would be supported by the cohesive party. This is even if the status quo is as far away as it is.
\subsubsection{Consensus Government: Compromise but Possible Gridlock}
\textbf{Assumptions:}
\begin{itemize}
	\item $A$ and $B$ are in the Left Party, $C$ is in the Center Party while $D$ and $E$ are in the Right Party
	\item The Left Party and Center Party are in coalition
	\item $B$ is the Prime Minister $\rightarrow$ \textbf{$B$ is the agenda setter}
	\item B has to make a compromise proposal because \textbf{C is a veto player}
\end{itemize}
\begin{center}
\begin{tikzpicture}
\coordinate [label=left:$Left$] (Left) at (-5,0);
\coordinate [label=right:$Right$] (Right) at (5,0);
\coordinate [label=below:$A_L$] (A) at (-3.5,0);
\coordinate [label=below:$B_L$] (B) at (-2,0);
\coordinate [label=below:$C_C$] (C) at (0,0);
\coordinate [label=below:$D_R$] (D) at (1,0);
\coordinate [label=below:$E_R$] (E) at (3,0);
\coordinate [label=above:$SQ$] (SQ) at (2, 1);
\coordinate [label=above:$X$] (X) at (-2, 1);
\draw [thick, <->] (-2, -.7) -- (2, -.7);
\draw [->] (SQ) -- (2, 0);
\draw [->] (X) -- (-2, 0);
\draw (Left) -- (Right);

% Draw in status quo with chosen policy X and winset from C equi-distant both sides to SQ


\end{tikzpicture}
\end{center}
Now with a veto player, there is a winset which is the set of policies that the veto player, $C$, prefers to the status quo. This is shown by the double-headed arrow interval between $SQ$ and $X$. If $B$ proposes anything in this interval, $C$ prefers it to $SQ$ and is likely to agree with B to support the coalition.\\

However, if $SQ$ is in between $B$ and $C$, there is likely to be gridlock. This is because $B$ wants policy to move towards $B$ while $C$ wants policy to move towards $C$. We can show this \textbf{gridlock interval} in the illustration below:
\begin{center}
\begin{tikzpicture}
\coordinate [label=left:$Left$] (Left) at (-5,0);
\coordinate [label=right:$Right$] (Right) at (5,0);
\coordinate [label=below:$A_L$] (A) at (-3.5,0);
\coordinate [label=below:$B_L$] (B) at (-2,0);
\coordinate [label=below:$C_C$] (C) at (0,0);
\coordinate [label=below:$D_R$] (D) at (1,0);
\coordinate [label=below:$E_R$] (E) at (3,0);
\coordinate [label=above:$SQ$] (SQ) at (-1, 1);
\coordinate [label=above:$X$] (X) at (-2, 1);
\draw [thick, <->] (-2, -.7) -- (0, -.7);
\draw [->] (SQ) -- (-1, 0);
\draw [->] (X) -- (-2, 0);
\draw (Left) -- (Right);

% Draw in status quo with chosen policy X and winset from C equi-distant both sides to SQ


\end{tikzpicture}
\end{center}
Simply, a proposal of a new policy $X$, by $B$ would be immediately vetoed but $C$ as it is further away from $C$ in the current status quo. The whole interval presents gridlock assuming that the parties are looking to get policy closest to their preference point.

This gridlock interval changes with ideological distance within coalitions and even opposing parties. The USA, with their parties now so far apart, shows the massive gridlock that they have in their system. Due to this, the amount of policy that can be passed is minimal.
\begin{center}
Tsebelis states: If an exogenous shock occurs, a government with many veto players with big ideological distances among them cannot handle the situation and cannot agree on the necessary policies. (2002 p.185)
\end{center}

\subsection{Summary of Veto Players and Agenda Setters}

\newpage
\section{Office vs Policy Seeking Theories of Coalitions}
\textbf{Definitions:}
\begin{itemize}
	\item \textbf{Office Seeking Theory (Riker 1962):} parties, when negotiating, will try to maximise the number of cabinet seats they can achieve, only \textit{minimum winning coalitions} should form as this minimises the amount of cabinet seats that have to be shared between coalition partners
	\item \textbf{Policy Seeking Theory (Axelrod 1970):} parties will try to maximise their influence over policy outcomes, only \textit{connected coalitions} should form between parties that are next to each other on a policy scale, as this minimises the likely policy disagreements between parties
\end{itemize}
\textbf{Note that these two things are not mutually exclusive and it is entirely possible to find minimum-connected-winning coalitions!}

\subsection{Office Seeking Theory}
The idea of the office seeking theory is to maximise the seats in office that you will take. This means that you will attempt to form a coalition that minimises the amount of portfolios that you need to distribute outside your own party. This comes from \textbf{Gamson's Law:}
\begin{center}
	\textbf{Gamson's Law} states that cabinet portfolios will be distributed among government parties in strict proportion to the number of seats that each party contributes to the government's legislative seat total i.e. the amount of seats your party has should proportionally translate to the amount of portfolios.
\end{center}
Thus the desire to maximise seats leads to the formation of a \textbf{minimal winning coalition (MWC).} This is when there is \textbf{just enough parties and no more} required to control a legislative majority. A \textbf{least minimal winning coalition} is a MWC with the lowest number of surplus seats. For example, if we have:
\begin{center}
\begin{tabular}{|c|c|c|}

\hline
Party & \% Seats & Party Position\\
\hline
Party 1 & 45\% & Far Right\\
\hline
Party 2 & 7\% & Far Left\\
\hline
Party 3 & 4\% & Center\\
\hline
Party 4 & 2\% & Center Left\\
\hline
Party 5 & 36\% & Center Right\\
\hline
Party 6 & 6\% & Moderate Right\\
\hline

\end{tabular}
\end{center}

In this, on the assumption that we need 51\% to control a majority, we can find a few coalitions such as Party 1 and Party 2 which gains 52\% of the seats, Party 2, 4, 5 and 6 which gives us 51\% and Party 1, 3 and 4 which also gives us 51\%. The coalitions which give us 51\% of the seat share are \textbf{least minimal winning coalitions.} The second assumption, after choosing the minimal winning coalition is that the parties would choose the least minimal so that the seat share is less distributed among the coalitions. Thus this is what we can conclude from a \textbf{purely office seeking world.} In this world, party position does not matter, as shown above.

\subsection{Policy Seeking Theory}
The idea of policy seeking theory is to create coalitions based on policy concessions. This is because parties seeking to be in power are aiming to implement their own policies that they are advocating for. They no longer simply seek office. Hence, when making concessions on policy, the positions of parties matter greatly. This means that you want to form governments with parties close to you in policy space. 

Political scientists often refer to this type of coalition as a \textbf{connected coalition} where members of the coalition are next to one another in the policy space. If we take a political spectrum as illustrated below:
\begin{center}
\begin{tikzpicture}
\coordinate [label=left:$Left$] (Left) at (-5,0);
\coordinate [label=right:$Right$] (Right) at (5,0);
\coordinate [label=below:$A_L$] (A) at (-3.5,0);
\coordinate [label=below:$B_L$] (B) at (-2,0);
\coordinate [label=below:$C_C$] (C) at (0,0);
\coordinate [label=below:$D_R$] (D) at (1,0);
\coordinate [label=below:$E_R$] (E) at (3,0);
\draw (Left) -- (Right);

% Draw in status quo with chosen policy X and winset from C equi-distant both sides to SQ

\end{tikzpicture}
\end{center}

We can see that Parties such as $A$ and $B$, if they control the majority of the vote, can form a connected coalition. However, if parties $A$ and $E$ were to form a coalition, this would not be connected since they are on opposite sides of the spectrum.

A second implication of the purely polic--seeking logic is that you will choose the connected least minimal winning coalition. This is to reduce the amount of policy concessions that you'd have to have to be in power. This means that we take our logic from the office seeking theory and combine it with policy seeking incentives to arrive at \textbf{the connected least minimal winning coalition.}

\subsection{Summary of Office vs Policy Seeking Coalitions}

\newpage
\section{Principal-Agent Theory of Independent Institutions}
\textbf{Definitions:}
\begin{itemize}
	\item \textbf{Principal:} the delegator, usually the politicians who outsource certain tasks to institutions
	\item \textbf{Agent:} the delegatee, usually the institution who acts on behalf of the politicians
	\item \textbf{Moral Hazard:} occurs when the agent has the opportunity to take agents that are hidden from the principal
	\item \textbf{Adverse Selection:} occurs when the agent has attributes that are hidden from the principal
	\item \textbf{Ex Ante Mechanism:} helps principals learn about their agents before these agents are chosen e.g. education and experience filtering
	\item \textbf{Ex Post Mechanism:} helps principals learn about agents actions after they have occurred. This is usually performance and how well they have conformed to the principal's ideas
	\begin{itemize}
		\item In a \textbf{Fire Alarm} system, the principal relies on iformation from others to learn about their agent.
		\item In a \textbf{Police Patrol} system, the principal monitors their agents actions
	\end{itemize}	 
\end{itemize}
This theory was first used in the study of US Government from a political science perspective where we saw legislature vs the executive. However, it is now used increasingly everywhere through analysis of parliamentary government, EU politics, international organisations etc.

\subsection{Principal Agent (PA) Theory Insights}
The main insights of this theory are as follows:
\begin{itemize}
	\item Explains why principals delegate to independent agents
	\item Explains why agents do not do as they are told i.e. \textbf{policy drift}
	\item Explains how policy drift can be reduced
\end{itemize}
These will be explained in the following subsections.
\subsection{Why Delegate?}
Here are some reasons why:
\begin{itemize}
	\item To \textbf{protect particular policies from short-term change} by a particular political majority
	\item To \textbf{establish a credible commitment} to a particular policy
	\item To \textbf{reduce workload} and enhance efficiency in decision-making
	\item To increase \textbf{the use of experts in policy making}
	\item To avoid taking blame for unpopular policies i.e. \textbf{blame shifting}
	\item To produce \textbf{stable and predictable policies}
\end{itemize}
In terms of Courts and Central Banks, here are some reasons to delegate:
\begin{center}
\begin{tabular}{c|c}
Courts & Central Banks\\
\hline
Protect Human Rights & Time inconsistent preferences\\
Complete legislative contracts & End political business cycle\\
Control other agencies & Lock-in stable macro-economic environment\\
\hline
\end{tabular}
\end{center}
Hence, delegation to these institutions ensure more stability in policy over time across the board whilst also providing credibility to their agreements

\subsection{Policy Drift}
Yet, despite all the delegation, policy drift, which is the difference between what the principal desires and what the agent actually does, can still occur. This is even with our mechanisms of selection. Here are reasons why:\\\\
\textbf{The Cause:} the agent does not have the same \textit{policy preferences} as the principal.\\\\
\textbf{Causes of Differences in Preferences:}
\begin{itemize}
	\item Appointment procedure $\rightarrow$ biased preferences of the agent such as judges appointing judges
	\item Specialised knowledge of agents that principals do not have
	\item Instituional/bugetary interests of agency that might not be aligned with the principals ideas
	\item Capture of agent by private interests $\rightarrow$ lobbying, for example, by big banks for the central bank to be \textbf{hawkish} or \textbf{dovish} in certain periods of the economic cycle
\end{itemize}
In terms of Courts and Central banks, here are some reasons for policy drift:
\begin{center}
\begin{adjustbox}{width=.85\textwidth}
\begin{tabular}{c|c}
Courts & Central Banks\\
\hline
Judges tend to be old, rich, white men & Central bankers are naturally hawks $\rightarrow$ against inflation \\
Liberal legal training $\rightarrow$ protect individuals vs the state & Epistemic community of neoclassical economists\\
Asymmetric access to courts $\rightarrow$ they favour the rich and powerful & Closer connection to financial institutions\\
\hline
\end{tabular}
\end{adjustbox}
\end{center}

\subsection{Controlling Policy Drift}
Here are a few methods of controlling policy drift:
\begin{itemize}
	\item Appoint new agents
	\item Restrict bugdet of an agency
	\item Write detailed legislation \textbf{to limit discretion of agents}
	\item Parliamentary scrutiny of agency actions
	\item Delegate to courts to constrain agency e.g. \textbf{competition authority}
	\item New laws e.g. \textbf{revise agency mandate}
\end{itemize}
In the context of Courts and Central banks:
\begin{center}
\begin{adjustbox}{width=.85\textwidth}
\begin{tabular}{c|c}

Courts & Central Banks\\
\hline
Politicised Supereme Courts & Partisan Appointment of Governors\\
Partisan Appointment of Judges & Appointment of non-economists to boards\\
Creation of Special Courts & Separation of mandate $\rightarrow$ politicians set inflation target, banks the interest\\
\hline

\end{tabular}
\end{adjustbox}
\end{center}



\subsection{Summary of the PA Theory}

\end{document}